---
title: "Tutorial"
editor_options: 
  chunk_output_type: console
bibliography: references.bib
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  compact = TRUE
)
```

```{r setup}
library(seaR)

# custom Functions for this analysis

is_homophily <- function(x) {
  ## 1 if group (size > 1) is homophilous
  ## 0 otherwise
  out <- str_extract_all(x, pattern = "M|F")
  as.integer(map_lgl(out, \(x) (all(x == "M") | all(x == "F")) & length(x) > 1))
}
```

## Data

To do: Add origin story of the data.

| Variable    | Type      | Description                                                                                                                                                                                                                                                              |
|---------------|---------------|------------------------------------------|
| `mr_id`     | numeric   | unique id                                                                                                                                                                                                                                                                |
| `teamtot`   | numeric   | number of people in the team                                                                                                                                                                                                                                             |
| `q217_*`    | character | gender of owner; may include category of non-person which refers to institutional owners (such as banks)                                                                                                                                                                 |
| `wtw1`      | double    | sampling weight                                                                                                                                                                                                                                                          |
| `men`       | numeric   | count of men in team                                                                                                                                                                                                                                                     |
| `women`     | numeric   | count of women in team                                                                                                                                                                                                                                                   |
| `allmale`   | numeric   | indicator variable                                                                                                                                                                                                                                                       |
| `allfemale` | numeric   | indicator variable                                                                                                                                                                                                                                                       |
| `spouse`    | numeric   | indicator; are there any spouses/cohabiting partners among the founding team members?                                                                                                                                                                                    |
| `autonomy`  | numeric   | An indicator which assesses the role of institutional owners and sponsors. The values are (1) no outside influence; (2) independent startup with institutional owners; (3) franchise or multilevel marketing enterprise; and (4) startup sponsored by existing business. |

```{r}
dplyr::glimpse(teams)
```

### Reshaping data

```{r, message=FALSE}
library(tidyverse)
theme_set(theme_light(base_family = "Optima"))
```

Step 1. Get the original data into the right shape (including the creation of an `.id` variable).

*Note the use of survey weights.*

```{r}
teams <- teams |> 
  ## Replace gender role names with first letter (i.e., M, F, N)
  mutate(across(starts_with("q217"), \(x) substr(x, start = 1, stop = 1))) |> 
  ## Ignore non-persons
  mutate(across(starts_with("q217"), \(x) ifelse(x == "N", NA, x))) |> 
  ## Create unique SE id
  mutate(.id = sort_collapse(q217_1, q217_2, q217_3, q217_4, q217_5)) |> 
  ## move id variable to beginning of data frame
  relocate(.id)

d <- teams |> 
  count(.id, spouse, wt = wtw1) 

print(d, n = Inf)
```

Step 2. Get the risk set.

```{r}
gender <- c("M", "F")
rs <- risk_set(1:5, gender) 
rs
```

The risk set is half the size as it should be because we are interested in whether there are spouses present in some of these groups or not.

```{r}
rs <- rs |> 
  mutate(spouse = list(c(0L, 1L))) |> 
  unnest(spouse)

glimpse(rs)
```

Step 3. Remove structural zeroes (i.e., groups of size 1 with spouse = 1).

```{r}
rs <- rs |> 
  filter(!(.size == 1 & spouse == 1)) 
```

Step 4. Merge the risk set with real data.

The risk set has 38 structural events, but we only have 28 of them in our data.

The idea is to merge them and have the missing values of `n` replaced with zeroes.

```{r}
d <- rs |> 
  full_join(d) |> 
  mutate(n = replace_na(n, 0L)) 
```

Step 5. Add additional `homophily` and `size_one` variables.

I also created a `y` variable that rounds up the weighted counts (`n`).

```{r}
d <- d |> 
  mutate(homophily = is_homophily(.id)) |> 
  mutate(y = round(n)) |>
  mutate(size_one = as.integer(.size == 1L)) 

d
```

Step 6. Calculate event probabilities

```{r}
d$lp <- event_probs(d, identities = c("M", "F"), log = TRUE)
attr(d$lp, "identities_prob") 
```

Step 7. The probabilities are not quite right because I have not taken into account the probabilities of there being a spouse or not in the group.

I haven't figured out a way to do this properly, but there should be a way to modify `event_probs` in the future to deal with these!

```{r}
d <- d |> 
  group_by(.size) |>
  mutate(lp = unclass(lp)) |> 
  mutate(lp = case_when(
    .size == 1 ~ lp,
    .size > 1 & spouse == 0 ~ lp + log(sum((spouse == 0) * n ) / sum(n)),
    .size > 1 & spouse == 1 ~ lp + log(sum((spouse == 1) * n ) / sum(n))
  )) |> 
  ungroup()

d |> ## sanity check, all conditional probabilities should sum up to one
  summarize(p = exp(log_sum_exp(lp)), .by = .size) 
```

### Expected counts

```{r}
d |> 
  mutate(expected = exp(lp) * sum(y), .by = .size) |> 
  select(.id, spouse, y, expected) |> 
  arrange(spouse) |> 
  gt::gt(rowname_col = ".id") |> 
  gt::fmt_number(decimals = 1) |> 
  gt::opt_table_font("Optima") 
```

### Models 1-4

The following code sort-of-replicates Table 4 in @ruef2003 [pp. 209]. They account for teams of size 4 or 5 with two spouses, which adds 11 more structural events.

```{r}
d <- d |> 
  mutate(size_category = factor(.size))

mod1 <- glm(y ~ size_category, data = d, family = poisson("log"), offset = lp)
mod2 <- update(mod1, y ~ .size)
mod3 <- update(mod1, y ~ size_one + .size + homophily)
mod4 <- update(mod1, y ~ size_one + .size + homophily*spouse)
```

```{r}
cm <- c(
  "(Intercept)" = "(Intercept)",
  ".size" = "Team Size",
  "size_one" = "Size Category [1]",
  "size_category2" = "Size Category [2]",
  "size_category3" = "Size Category [3]",
  "size_category4" = "Size Category [4]",
  "size_category5" = "Size Category [5]",
  "homophily" = "Gender Homophily",
  "homophily:spouse" = "Partners/spouse × \nGender Homophily"
)

modelsummary::msummary(
  list("Model 1" = mod1, "Model 2" = mod2, "Model 3" = mod3, "Model 4" = mod4), 
  coef_map = cm,
  stars = TRUE,
  gof_map = c("nobs", "aic", "bic"),
  output = "gt"
) |> gt::opt_table_font("Optima")
```

## References 

