---
title: "Definitions"
bibliography: references.bib
csl: ajs.csl
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE}
library(seaR)
library(dplyr)
```

## Structural Event

The term "structure" means *some combination* of basic actors (or identities or roles), with no strict bound on the number of those actors. The term "event" is used with the same meaning that a random event has in probability theory---i.e., some well-defined *subset* of a *sample space.*

Suppose we have two gender roles (`F`, `M`), three occupational roles (`A`, `B`, `C`), and group sizes ranging from 1 to 10.

**Is there an upper limit?**

```{r}
gender <- c("F", "M")
occupation <- c("A", "B", "C")

risk_set_size(1:5, gender, occupation)
```

## Risk Set

The risk set $s(H)$ of a structural event enumerates all possible combinations over a set of roles $(N)$, subject to group size $(r)$ and restrictions over permissible role combinations.

The number of combinations for a multiset of N roles is calculated as:

$$
s(H) = \sum_{r = 1}^{r(H)} {r + |N| -1 \choose r } = \sum_{r = 1}^{r(H)} \frac{(r + |N| -1)!}{r! (|N| -1)!}
$$

*Note. This is simply the standard way of calculating the number of possible samples of size* $r$ *out of a population of size* $|N|$*, assuming we are looking for unordered samples with replacement, and then summing over all possible values of* $r$.

For example, for two gender roles (`F`, `M`) and four occupational roles (`A`, `B`, `C`, `D`), we have the following risk set:

```{r}
gender <- c("F", "M")
occupation <- c("A", "B", "C")
rs <- risk_set(1:2, gender, occupation)
glimpse(rs)
```

```{r}
tibble::deframe(rs)
```

## Structural Zeroes

"If there are a priori restrictions imposed on role combinations (for instance, if women in a given society are not allowed to hold certain occupations), then the role set must be reduced accordingly" [@ruef2003, pp. 219]. We can do this by filtering the `rs` data frame.

For example, if it happens to be the case that `M` is not legally allowed to have a `C` occupation:

```{r}
rs_restricted <- rs |>
  filter(!grepl(.id, pattern = "MC"))
  
rs_restricted
```

## Event Probability

If structural events are drawn from a population of entrepreneurs that is 40 percent manual, 30 percent service, and 30 percent professional, then the expected probability of obtaining a three member founding team with one manual worker and two service sector workers under an assumption of statistical independence is $\Pr(E \mid 3) = \frac{3!}{2! \times 1!} (.4^1 \times .3^2 \times .3^0) = .108$. The event probability reflects the fact that there are three different ways to draw the participants. By comparison, the probability of obtaining a three-member team that consists only of manual workers is $\Pr(E \mid 3) = \frac{3!}{3!} (.4^3 \times .3^0 \times .3^0) = .064$.

```{r}
probs <- c("M" = .4, "S" = .3, "P" = .3)

dmultinom(x = c(1, 2, 0), prob = probs)
dmultinom(x = c(3, 0, 0), prob = probs)
```

## Code
